---
title: "Practical Machine Learning Course Project"
author: "datakatsudon"
date: "September 26, 2015"
output: html_document
---

#Loading Packages and Datasets:
```{r, echo=TRUE}
library(caret)
library(randomForest)
set.seed(123)
trainingset <- read.csv("pml-training.csv")
testset <- read.csv("pml-testing.csv")
```

#Cleaning up the Data:
From the trainingset I read, I created a new dataframe (newaccel) that has accelerometer readings without NA values in the columns:
```{r, echo=TRUE}
accel <- trainingset[,c("user_name", colnames(trainingset)[grep("accel", colnames(trainingset))], "classe" )]
  ##subsetted the user_name column, classe column, and columns that have accelerometer readings
rmtest <- apply(accel, 2, function(x) length(which(is.na(x))))
  ##returns the number of NAs in each column
newaccel <- accel[, c(1,2,4,5,6,7,9,10,11,12,14,15,16,17,19,20,21,22)]
  ##created a new dataframe that consists of columns that don't have NA values

acceltest <- testset[,c("user_name", colnames(testset)[grep("accel", colnames(testset))])]
nacceltest <- acceltest[, c(1,2,4,5,6,7,9,10,11,12,14,15,16,17,19,20,21)]
  ##subsetted same columns as newaccel dataframe, but did not include column 22 (classe)
  ##as there is no classe column in this dataframe
```

#Creating the Model:
I passed the newaccel dataframe through the function train() and passed the cross-validate argument through the trControl argument:
```{r, echo=TRUE}
trControl <- trainControl(method="cv", number=5)
testfit <- train(classe~., data=newaccel, method="rf", trControl=trControl)
```
According to the following information taken from my testfit model, the estimate of error rate is 4.65%. This is about one error per 20 observations. 
```{r, echo=TRUE}
testfit$finalModel
```

I predicted the classe letter of observations in my newaccel dataframe (from the training data) based on my model testfit and took a look at the statistics using the confusionMatrix() function:
```{r, echo=TRUE}
valpred <- predict(testfit, newaccel)
valerror <- confusionMatrix(valpred, newaccel$classe)
valerror
```

##Prediction Results:
I applied the model to my nacceltest dataframe that was originally missing the classe column and predicted the classe letter:
```{r, echo=TRUE}
finalprediction <- predict(testfit, nacceltest)
finalprediction
```
